### 2012-04-18 4.1.1.tex
chapters 2 and 4 have been swapped

% 4.1.1: Gradience in judgements

at least as far back as \emph{LSLT} \citep{LSLT}

%and has become a recent feature of the discussion
is that non-gradient models are totally incompatible with psycholinguistic results.

\subsubsection{Task model}

\citet[][16]{Schutze1996}

\citet{Schutze2005}
\citet{Schutze2011}

%\citet{Coleman1996}

\subsubsection{Falsifiability}

\citet{Armstrong1983}

% wordlikeness critique
\citet{Yang2008a}

a mix of ``declarative'' knowledge,
corresponding to rote facts and thus the lexicon,
and ``procedural'' knowledge, or knowledge of generative procedures 
(in the sense of J. \citealt{Anderson1993}).

% math cog
\citet{Logan1988}
\citet{Sfard1991}
\citet{Delazer2005}

>>>>

\subsection{Data sources}

The English CELEX database \citep{CELEX},

\citet{Aronoff1976}

\citet{Harley2009}

\citet{Taft1975} ?
\citet{Taft1981} ?
\citet{Taft2004a} ?

\citet{Halle1962}

% 2.1.4: k

\subsubsection{Corpus counts}

Under the phonological analyses described in the next two sections, the 6,876 simplex words in CELEX contain 21 unique medial codas and 40 unique medial onsets. Of these 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, for a 18.8\% saturation rate.

### 2012-04-18 4.1.2.tex
More arranging.
chapters 2 and 4 have been swapped

% 4.1.2: The ontology of wordlikeness

\subsubsection{Gross phonotactic violation}

position in \citet{SPE}

\citet{Shibatani1972}
\citet{Hooper1973}

peripheral \citep[in the sense of][81]{LGB}

As an example of this applied ot phonotactics, 
\citet{Borowsky1989} assumes that there are peripheral morph-internal syllable contact clusters.

\subsubsection{Phonological statistics}

\citet{SPE}

This statement is made very strongly by \citet{Coleman1996}

\subsubsection{Lexical similarity}

\citet{LSLT}

\citet{Greenberg1964}
\citet{Ohala1986b}

A recent, very articulated model is presented by \citet{Bailey2001}.

% simple model
This is simply because more complex models fail to

% similarity vs. rules
% as a final note, \citet{Hahn1998}

>>>>

\subsection{Syllabification}

It is necessary to split medial consonant clusters into the codas and onsets that make them up. While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic. For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language. Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which contrast \emph{chemistry} and \emph{ministry}.

An automated syllabification system was constructed and applied to medial clusters in the CELEX data. This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.  There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification. The technique used is a variation on the theme of onset maximization \citep[42f.]{Kahn1976}, which parses word-medial clusters so that the onset is as large as possible. Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty. In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.

There is one case where unchecked maximization of medial onsets produces incorrect syllabifications. When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. Onset maximization makes the right predictions for words like \emph{propaganda}, where the medial cluster cannot be an onset, but \citet[][55]{Harris1994} notes that it makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty. The approach adopted here is to first apply onset maximization, then to reassign the first consonant of a medial onset to the preceding coda if it is immediately preceded by a stressed lax vowel.

If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables. For this reason, it is assumed here that affricates are single segments for the purposes of syllabification. This assumption is motivated by the fact that affricates pattern with simple segments in many languages; for instance, the complex onsets in Classical Nahua are affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.

%In fact, [t.ʃ, d.ʒ] clusters are not found in simplex English words, despite the fact that that affricates occur as medial onsets in clusters (e.g., \emph{tru}[n.tʃ]\emph{eon}, \emph{so}[l.dʒ]\emph{er}).

### 2012-04-18 4.1.3.tex
chapters 2 and 4 have been swapped

The syllables derived by the above procedure must further be divided into onset, nucleus, and coda, and mapped to underlying representation. In most cases, this is trivial: the nucleus is a pure vowel and can be easily separated from the consonants, which surface faithfully. However, the status of several surface phones in the syllable and the relationship of surface segments to underlying forms is not always obvious. The treatment of those segments is described and motivated below.

@@

There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}. The strongest piece of evidence for the pure allophonic analysis (and an attentant process of /g/-\textsc{Deletion}) is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is adopted here and formalized below in Section \ref{cnpasection}.
%\citealt[][62]{Halle1985a}),

@@

I assume that the [j] onglide in words such as \emph{ass}[j]\emph{ume} is present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}). The primary motivation for this assumption is the observation that presence or absence of this glide is contrastive (e.g. \emph{coup} $\sim$ \emph{queue}, \emph{booty} $\sim$ \emph{beauty}). I further assume that the front glide is not part of the onset except in the case where the onset would ortherwise be null (e.g., \emph{jun}[j]\emph{or}).
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}. However, when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm], and speakers judge other following vowels to be anomalous in nonce words \citep{Moreland2009}. This defective distribution of vowels following [j] follows if the onglide in this context is the first component of a phonological diphthong, and thus part of the nucleus (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear. There is a great deal of external support for this position as well. The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003}, to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

\ex VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}: \\
\begin{tabular}{l l l}
a. & kəræb  & `crab'  \\
   & bəlid  & `bleed' \\
b. & kəwin  & `queen' \\
   & kəwoʊt & `quote' \\
c. & kut    & `cute'  \\
   & musɪk  & `music  \\ 
\end{tabular}
\xe

\noindent
VBR breaks up complex onsets, including back onglide clusters in (\nextx b). However, in (\nextx c), which contain the front onglide in standard English, no epenthesis occurs: the glide is simply absent. This also suggests that the front onglide is part of the onset, not the nucleus.

@@

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [ju\lm], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}.

@@

Like this study, \citet{Pierrehumbert1994} uses a dictionary that transcribes RP, in which word-medial post-vocalic \emph{r} has been lost. There are independent reasons to believe that \emph{r}-full dialects of English assign post-vocalic \emph{r} to the nucleus and is not part of syllable contact clusters in such dialects. First, \citet[][255]{Harris1994} notes that the number of vowel contrasts is greatly reduced before \emph{r} when compared to other coronal consonants. For instance, the majority of North America has lost the historical contrast between \emph{Mary}, \emph{marry}, and \emph{merry}, and one holdout, Philadelphia, is losing the the contrast between \emph{merry} and \emph{Murray} \citep[14f.]{ANAE}. Another source of evidence for the nuclear status of post-vocalic \emph{r} comes from its failure to trigger two processes in English variable phonology. \citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.

\ex /t/-\textsc{Glottalization} in British English \citep[after][195, 258]{Harris1994}: \\
\begin{tabular}{l l l@{} l}
a. & fis[t]   & * & fis[ʔ]   \\
   & mis[t]er & * & mis[ʔ]er \\
b. & par[t]   &   & par[ʔ]   \\
   & car[t]on &   & car[ʔ]on \\
\end{tabular}
\xe

@@

\ex /t, d/-\textsc{Deletion} in American English: \\
\begin{tabular}{l l l@{} l}
a. & be[lt]  &   & be[l]  \\
%   & we[ld]  &   & we[l]  \\
%   & cha[nt] &   & cha[n] \\
   & me[nd]  &   & me[n]  \\
%b. & fl[ɜ˞]  & * & fl[ɜ˞]  \\
%b. & fl[ɝt]  & * & fl[ɝ]  \\
b. & sh[ɚt]  & * & sh[ɚ] \\
%   & w[ɚd]   & * & w[ɚ]  \\
   & c[ɚd]   & * & c[ɚ]  \\
\end{tabular}
\xe

>>>>

To identify the components of syllable contact clusters, the syllables identified by the above procedure must be further separated into onset, nucleus, and coda, and mapped to underlying representation. This is trivial in most cases: the nucleus is easily separated from onset and coda and surface consonants are faithful to underlying representation. The treatment of the few difficult cases is described below. Ambiguous segments are assigned to onset, nucleus, or coda, and to underlying form, using a distributional heuristic.

@@

There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right, as demanded by Kiparsky's Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ \citep[][65]{Borowsky1986}. The strongest piece of evidence for treating the velar nasal as a pure allophone is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here. English nasal place allophony is formalized in Section \ref{cnpasection} below.
%\citealt[][62]{Halle1985a}),

@@

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). 
%\footnote{English glides are transcribed here as full segments, not as ``subsegments'', as this distinction does not appear to be meaningful for English, or empirically motivated for other languages \citep{Rubach2002}.}
There is extensive motivation for this principle. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [u\lm]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [u\lm] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [ju\lm] also suggests the glide is nuclear. There is also strong external support for this principle. The [ju\lm] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [ju\lm] at the expense of the onset is also found in speech errors, e.g., [kju\lm]\emph{mor homponent} for [hju\lm]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}. Finally, \citet{Buchwald2005} considers [j] onglides in the speech of VBR, an aphasic patient who has difficulties producing complex onsets.

\begin{example}
VBR's complex onsets \citep[79--80, his transcriptions]{Buchwald2005}:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l}
a. & kəræb  & `crab'  \\
   & bəlid  & `bleed' \\
b. & kəwin  & `queen' \\
   & kəwoʊt & `quote' \\
c. & kut    & `cute'  \\
   & musɪk  & `music  \\ 
\end{tabular}
\end{example}

\noindent
VBR breaks up complex onsets with epenthesis, including those that consist of a consonant and a back onglide cluster (\lastx b). However, no epenthesis occurs between a consonant and a front onglide; rather, the glide is absent (\lastx c). The failure of the front onglide to pattern with other consonant clusters suggests once again that the glide is part of the nucleus.

@@

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [ju\lm], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}. And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.

@@

Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost. In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters. Many vowel contrasts are suspended before \emph{r} \citep[][255]{Harris1994}; for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}. Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants. \citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.

\begin{example}
/t/-\textsc{Glottalization} in British English \citep[after][195, 258]{Harris1994}:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l@{} l}
a. & fis[t]   & * & fis[ʔ]   \\
   & mis[t]er & * & mis[ʔ]er \\
b. & par[t]   &   & par[ʔ]   \\
   & car[t]on &   & car[ʔ]on \\
\end{tabular}
\end{example}

@@

\begin{example}
/t, d/-\textsc{Deletion} in American English:

\vspace{0.5\baselineskip}
\begin{tabular}{l l l@{} l}
a. & be[lt]  &   & be[l]  \\
%   & we[ld]  &   & we[l]  \\
%   & cha[nt] &   & cha[n] \\
   & me[nd]  &   & me[n]  \\
%b. & fl[ɜ˞]  & * & fl[ɜ˞]  \\
%b. & fl[ɝt]  & * & fl[ɝ]  \\
b. & sh[ɚt]  & * & sh[ɚ] \\
%   & w[ɚd]   & * & w[ɚ]  \\
   & c[ɚd]   & * & c[ɚ]  \\
\end{tabular}
\end{example}
